---
title: "Final Project"
author: "miranda_moe"
date: "2026-05-09"
output: 
  github_document:
    toc: true
    toc_depth: 2
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this project, we will if we can build a classification model to learn features of model descriptions to predict genre it represents. We will use a dataset from Kaggle that has various characteristics of movies grouped in different genres. Each genre has its own csv, and there are 16 genres in the dataset in total. The following is the link to original dataset:

https://www.kaggle.com/datasets/rajugc/imdb-movies-dataset-based-on-genre/data


## Packages

Load the packages we need for this project.

```{r load packages}
library(tidymodels)
library(tm)
library(ggplot2)
library(tidyverse)
library(modeldata)
library(rpart.plot)
library(tidytext)
library(stringr)
library(tidyr)
library(caret)
library(tidytext)
library(lubridate)
library(tibble)
```

## Data Frames

Now, we will load the csvs. 
```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
