---
title: "Final Project"
author: "miranda_moe"
date: "2026-05-09"
output: 
  github_document:
    toc: true
    toc_depth: 2
---

In this project, we will if we can build a classification model to learn features of model descriptions to predict genre it represents. We will use a dataset from Kaggle that has various characteristics of movies grouped in different genres. Each genre has its own csv, and there are 16 genres in the dataset in total. The following is the link to original dataset:

https://www.kaggle.com/datasets/rajugc/imdb-movies-dataset-based-on-genre/data


## Packages

Load the packages we need for this project.

```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidymodels)
library(tm)
library(ggplot2)
library(tidyverse)
library(modeldata)
library(rpart.plot)
library(tidytext)
library(stringr)
library(tidyr)
library(caret)
library(tidytext)
library(lubridate)
library(tibble)
library(textstem)
```

## Data Frames

Now, we will load the csvs. First, we would need to load every file that ends with csv. 

```{r pressure, echo=TRUE}
file_list <- list.files(pattern = "*.csv")
file_list
```

From raw data exploration, I have noticed that some movies have multiple genre listed. To resolve the issue of multiple labeling, I only kept movies where the CSV's genre appeared somewhere in the movie's genre list. For example, a movie with genres listed as "Animation, Adventure, Comedy" would be kept in animation.csv but excluded from adventure.csv if "Adventure" was not the intended label. This ensures each movie is only included when it genuinely belongs to that genre category.
To combine all 16 CSVs into a single dataframe, I used lapply() to loop through each file. For each file, I extracted the genre label from the filename by removing the .csv extension using gsub(), filtered rows using str_detect() to keep only movies matching that genre, added a new column final_genre to store the label, and selected only the relevant columns (movie_id, movie_name, description, final_genre). All the movies are now binded into one dataframe. 

```{r bind dataframes, message=FALSE}
# In scifi.csv, the genres are listed as "Sci-Fi". In sports.csv, they are listed as "Sport". so we would need to map that. 
genre_map <- c(
  "scifi" = "sci-fi",
  "sports" = "sport"
)
movies_list <- lapply(file_list, function(f) {
  genre_name <- gsub(".csv", "", f)
  search_term <- ifelse(genre_name %in% names(genre_map), genre_map[genre_name], genre_name)
  df <- read_csv(f) |>
    filter(str_detect(str_to_lower(genre), search_term)) |>
    mutate(final_genre = genre_name) |>
    select(movie_id, movie_name, description, final_genre)
  df
})

movies <- bind_rows(movies_list)
head(movies, 10)
```

## Text cleaning 

When I look through the new dataframe "movies", I noticed that some movies are missing descriptions (which happens across genres.) Those missing movie descriptions are indicated as "Add a Plot" rather than NA or blank cells. Therefore, we will remove those movies so it wouldn't mess up the classification. 

```{r}
filtered_movies <- movies %>%
  filter(!grepl("Add a Plot", description))
```

Here, we will do some basic text analysis, we will do that in the following order:

- tokenize
- remove numbers (mostly just years, such as 2018, 2019 etc)
- remove stopwords
- lemmatize (reduce words to stem. for example, 'playing' would become 'play')
- TF-IDF (high tf-idf means words that appear a lot in one movide but rarely across others)
- Document-Term Matrix (DTM) that will transform into a numeric matrix 

Before starting any text operations, I have noticed that some descriptions don't have the full descriptions in the original data and at the end, it would say " See full summary »". And, we also have parantheses in descriptions that would include years (for example, spinoff from a movie name (2018), or a character name with an action name in parantheses). I assumed those weren't super necessary for the model to learn the features. And, there are periods between words such as L.A. or C.I.A, and tokenization might split them, which might not result in a meaningful result. There are also single and double quotation marks around names, which we will also get rid of. Therefore, in the following steps we will remove those by using stringr functions. 

```{r}
movies_cleaned_descriptions <- filtered_movies %>%
  mutate(description = str_remove(description, "See full summary.*$"),
         description = str_remove_all(description, "\\s*\\([^\\)]+\\)"), #remove paratheses
         description = str_remove_all(description, "(?<=[A-Za-z])\\.(?=[A-Za-z])"), #remove periods
         description = str_remove_all(description, "[\"']")) #remove the single and double quotations
```


I've noticed that some movies still appeared across genres, since a movie that is labeled as 'Action, Adventure.' would appear in both csvs. And, since I will be using single label classification, I decided to only keep the first occurance of each movie. 

```{r}
movies_cleaned_descriptions <- movies_cleaned_descriptions %>%
  distinct(movie_id, .keep_all = TRUE)
```


Now that we've cleaned those, we will start text cleaning by using tidytext.

First step is tokenizing, removing stopwords and numbers. We will also convert the dataframe to tibbles before lemmatizing because tidyverse handles tibbles better and it will give us one token per row. 

```{r text cleaning with tidytext, message=FALSE}
text_movies <- movies_cleaned_descriptions %>%
  unnest_tokens(word, description) %>%
  filter(!grepl('[0-9]', word)) %>%
  anti_join(stop_words) 
  
lemmatized_movies <- tibble(text_movies) %>%
  mutate(word = lemmatize_words(word)) %>% #now we will lemmatize the words
  count(movie_id, movie_name, word, final_genre, sort = TRUE) %>%
  bind_tf_idf(movie_id, word, n) %>%
  arrange(desc(tf_idf))
lemmatized_movies
```


As a final step before building a model, we coverted the TF-IDF weighted tokens into a Document-Term Matrix using cast_dm(). A DTM represents each movie as a row and each unique word as a column, where values will be TF-IDF we calculated above. As mentioned before, DMF will covert the current one-token-per-row structure to numerical matrix. TF-IDF weighting was used instead of raw word counts so that common words appearing across many movies are already weighted, allowing the model to focus on words that are more distinctive and informative for predicting genre.

```{r document-term matrix, message=FALSE}
document_term_matrix_movies <- lemmatized_movies %>%
  cast_dtm(movie_id, word, tf_idf)
document_term_matrix_movies
```
We have currently 96045 unique words as columns, which will cause a problem when we convert to dataframe to train data. Therefore, we will remove the words that appear in very few movies.
```{r}
dtm_reduced <- removeSparseTerms(document_term_matrix_movies, 0.99)
dtm_reduced
```
We removed terms that are missing from more than 99% of documents. As a result, our novel terms came down to 171.

Now that we have a summary of DocumentTerm Matrix, we would need the categorical y (the genres) that we want to predict by using the DTM values. Therefore, we will create the outcome labels.

```{r outcomes, message=FALSE}
genre_labels <- lemmatized_movies %>%
  distinct(movie_id, final_genre)
head(genre_labels)
```


## Prepping Data Frames to build classification model

We would first need the dtm_reduce to be in a dataframe where we can join the outcome categories (labels). 

```{r dataframe, message=FALSE}
dtm_df <- as.data.frame(as.matrix(dtm_reduced))
head(dtm_df, 2)
```

As we can see above, movie ids are not a column and they appear as row names. We will first convert them to column called movie_id before we join the labels.

```{r join with labels, message=FALSE}
dtm_df$movie_id <- rownames(dtm_df)
head(dtm_df$movie_id)

dtm_df <- dtm_df %>%
  left_join(genre_labels, by = "movie_id")
head(dtm_df, 2)
```

Now that we've prepared the dataframe, we will proceed with building the model. 

## Building Classification Model
### Identifying the outcome
First, we will change the outcome variable into a factor

```{r factor, message=FALSE}
str(dtm_df$final_genre)
dtm_df$final_genre <- factor(dtm_df$final_genre)
levels(dtm_df$final_genre)
```

The outcome variable has 16 levels. Therefore, it is a multiclass classification problem.

### Examine the class balance

```{r}
table(dtm_df$final_genre)
```

Now, we will create a bar graph for genres

```{r}
ggplot(dtm_df, aes(x = final_genre, fill = factor(final_genre))) +
  geom_bar() +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    title = "Bar Graphs for Different Genres",
    ylab = "Count",
    fill = "Genres of movies"
  )
```

As we can see in the bar graph and table above, this dataset is not balanced. When class imbalance happens, the accuracy could be misleading because the common class (such as action) would be more represented than other classes during data training. Therefore, the results might skew towards the class with more data. 

### Create a train/test split

We will use an 80/20 split and set seed to 123. We will also use stratification so that the class imbalance is reserved in both training and testing sets.

```{r}
set.seed(123)
genre_split <- initial_split(dtm_df, prop = 0.8, strata = final_genre)
train_data <- training(genre_split)
test_data <- testing(genre_split)
```

