---
title: "Final Project"
author: "miranda_moe"
date: "2026-05-09"
output: 
  github_document:
    toc: true
    toc_depth: 2
---

In this project, we will if we can build a classification model to learn features of model descriptions to predict genre it represents. We will use a dataset from Kaggle that has various characteristics of movies grouped in different genres. Each genre has its own csv, and there are 16 genres in the dataset in total. The following is the link to original dataset:

https://www.kaggle.com/datasets/rajugc/imdb-movies-dataset-based-on-genre/data


## Packages

Load the packages we need for this project.

```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidymodels)
library(tm)
library(ggplot2)
library(tidyverse)
library(modeldata)
library(rpart.plot)
library(tidytext)
library(stringr)
library(tidyr)
library(caret)
library(tidytext)
library(lubridate)
library(tibble)
```

## Data Frames

Now, we will load the csvs. First, we would need to load every file that ends with csv. 

```{r pressure, echo=TRUE}
file_list <- list.files(pattern = "*.csv")
file_list
```

From raw data exploration, I have noticed that some movies have multiple genre listed. To resolve the issue of multiple labeling, I only kept movies where the CSV's genre appeared somewhere in the movie's genre list. For example, a movie with genres listed as "Animation, Adventure, Comedy" would be kept in animation.csv but excluded from adventure.csv if "Adventure" was not the intended label. This ensures each movie is only included when it genuinely belongs to that genre category.
To combine all 16 CSVs into a single dataframe, I used lapply() to loop through each file. For each file, I extracted the genre label from the filename by removing the .csv extension using gsub(), filtered rows using str_detect() to keep only movies matching that genre, added a new column final_genre to store the label, and selected only the relevant columns (movie_id, movie_name, description, final_genre). All the movies are now binded into one dataframe. 

```{r bind dataframes, message=FALSE}
movies_list <- lapply(file_list, function(f) {
  genre_name <- gsub(".csv", "", f)
  df <- read_csv(f) |>
    filter(str_detect(str_to_lower(genre), genre_name)) |>
    mutate(final_genre = genre_name) |>
    select(movie_id, movie_name, description, final_genre)
  df
})

movies <- bind_rows(movies_list)
head(movies, 10)
```

## Text cleaning 

When I look through the new dataframe "movies", I noticed that some movies are missing descriptions (which happens across genres.) Those missing movie descriptions are indicated as "Add a Plot" rather than NA or blank cells. Therefore, we will remove those movies so it wouldn't mess up the classification. 

```{r}
filtered_movies <- movies %>%
  filter(!grepl("Add a Plot", description))
```

Here, we will do some basic text analysis, we will do that in the following order:

- tokenize
- remove numbers (mostly just years, such as 2018, 2019 etc)
- remove stopwords
- lemmatize
- TF-IDF

Before starting any text operations, I have noticed that some descriptions don't have the full descriptions in the original data and at the end, it would say " See full summary »". And, we also have parantheses in descriptions that would include years (for example, spinoff from a movie name (2018), or a character name with an action name in parantheses). I assumed those weren't super necessary for the model to learn the features. And, there are periods between words such as L.A. or C.I.A, and tokenization might split them, which might not result in a meaningful result. There are also single and double quotation marks around names, which we will also get rid of. Therefore, in the following steps we will remove those by using stringr functions. 

```{r}
movies_cleaned_descriptions <- filtered_movies %>%
  mutate(description = str_remove(description, "See full summary.*$"),
         description = str_remove_all(description, "\\s*\\([^\\)]+\\)"), #remove paratheses
         description = str_remove_all(description, "(?<=[A-Za-z])\\.(?=[A-Za-z])"), #remove periods
         description = str_remove_all(description, "[\"']")) #remove the single and double quotations
```


Now that we've cleaned those, we will start text cleaning by using tidytext.

First step is tokenizing. 

```{r text cleaning with tidytext, message=FALSE}
text_movies <- movies_cleaned_descriptions %>%
  unnest_tokens(word, description) %>%
  filter(!grepl('[0-9]', word)) %>%
  anti_join(stop_words) 
  
tibble(text_movies) %>%
  print(n = 10)
```





